Adaptive evolution with aneuploidy and mutation

### 2022-03-04 ms/ms.tex
removed information prior that was not converged, because there were bug in it execution

Compared to other priors we tested, this prior produced lower WAIC, better posterior prediction plots, and more stable parameter estimates, as follows. %TODO IK what do you mean by better posterior prediction and is more stable parameter estimates still rellevant?
We tried to use additional growth curves. We used growth curves of \eumt\; (\emph{refined} strain from \citet{Yona2012}) and \anwt\; in \SI{39}{\celsius} to estimate $w_{2n*}/w_{2n+1}$ ~(\cref{fig:growth-curves}-L). The same prior was used for $w_{2n*}/w_{2n+1*}$. This prior resulted in WAIC 71.91, compared to 67.87 with the above informative prior. % generated with fixed-mutation-rate-and-WAICs.ipynb TODO IK update the WAIC with 1000 reps run
We also tried to use growth curves of \anwt\; and \euwt\; in \SI{30}{\celsius} to estimate $1-c$ ~(\cref{fig:growth-curves}-H). This estimation assumes that the cost of aneuploidy is the same in \SI{39}{\celsius} and \SI{30}{\celsius}; this might be incorrect, but we only assumed this to generate a prior distribution for the fitness values. The prior for $b$ was taken the same as for $c$. This prior did not converge: no parameter sets were found with an approximate likelihood greater than zero. %TODO IK check that this model had no bugs
% TODO IK it makes sense because bla bla (to explain)
Finally, we also tested a uninformative uniform prior with $\mathit{U}(1,6)$, for (i) all $w_{2n+1}$, $w_{2n+1^*}$, $w_{2n^*}$ and (ii) only for $w_{2n+1^*}$, $w_{2n^*}$, using the above informative prior for $w_{2n+1}$. In both cases the algorithm failed to converge.

For the mutation rate, $\mu$, and aneuploidy rate, $\delta$, we used uninformative uniform priors, $\mu \sim \mathit{U}(10^{-9}\ ,\ 10^{-5})$ and $\delta \sim \mathit{U}(10^{-6} , 10^{-2})$. A wider mutation rate prior, $\mu \sim \mathit{U}(10^{-9}\ ,\ 10^{-3})$, produced similar results.
% TODO IK, maybe wider mutation rate should be tried on a model without aneuploidy
%For a model without aneuploidy ($\delta=0$), we used a wider uniform prior for the mutation rate, $\mu \sim \mathit{U}(10^{-10},10^{-5})$.

@@

% TODO IK we are jumping between b c s and w2n* w2n+1 w2n+1*. we should report all models and results either with b c s or with w. you can compute b c s from w before computing the HDIs.
This estimated aneuploidy rate agrees with previous estimates. % TODO refs (9.7*10-5 in Zhu 2014, so it's lower). YR: we need more than one ref; also what is the units of Zho 2014?
The mutation rate corresponds to a mutation target size of $10^{4}$, assuming the mutation rate per base pair is roughly $2\cdot10^{-10}$~\citep{Zhu2014} or $3.3\cdot10^{-10}$~\citep{Lynch2008}. % TODO IK please add lynch2008 to bib file

@@

% Fig Curveball 
\begin{figure}[h]
	\centering
	% generated with growth_curves.ipynb and deg = '39' 
	\includegraphics[width=.95\textwidth]{../figures/evo39_fitness_39deg.pdf} 
	% generated with growth_curves.ipynb and deg = '30' 
	\includegraphics[width=0.95\textwidth]{../figures/evo39_fitness_30deg.pdf} 
	% generated with growth_curves_refined.ipynb and deg = '30' 
	\includegraphics[width=0.95\textwidth]{../figures/refined_vs_evo39_fitness_39deg.pdf}
\caption{
    \textbf{Fitness estimation from growth curves.}
    % TODO IK caption should explain what each panel shows.
    \textbf{(A-D)} Fitness estimation from growth curves of \euwt\ and \anwt\ at \SI{39}{\celsius}: $\hat w_{2n+1}/w_{2n}$=1.024 (95\% CI: 0.959 - 1.115).
    \textbf{(E-H)} Fitness estimation from growth curves of \euwt\ and \anwt\ at \SI{30}{\celsius}: $\hat w_{2n+1}/w_{2n}$= 0.986 (95\% CI: 0.979 - 0.994).
    \textbf{(I-L)} Fitness estimation from growth curves of \anwt\ and \eumt\ at \SI{39}{\celsius}: $\hat w_{2n*}/w_{2n+1}$=1.033 (95\% CI: 1.027 - 1.041).\\
    Growth curves previously described in \citet[Figs. 3C, 4A, and S2]{Yona2012}.
	Fitness estimated from growth curves using \texttt{Curveball}, a method for predicting results of competition experiments from growth curve data~\citep[\href{https://curveball.yoavram.com}{curveball.yoavram.com}]{Ram2019}. See \emph{Models and Methods, Prior distributions} for more details.
} 
\label{fig:growth-curves}
\end{figure}

>>>>

Compared to other priors we tested, this prior produced lower WAIC, better posterior prediction plots, and more stable parameter estimates, as follows. %TODO IK what do you mean by better posterior prediction and is more stable parameter estimates still rellevant?
We tried to use additional growth curves. We used growth curves of \eumt\; (\emph{refined} strain from \citet{Yona2012}) and \anwt\; in \SI{39}{\celsius} to estimate $w_{2n*}/w_{2n+1}$ ~(\cref{fig:growth-curves}-L). The same prior was used for $w_{2n*}/w_{2n+1*}$. This prior resulted in WAIC 71.91, compared to 67.87 with the above informative prior. % generated with fixed-mutation-rate-and-WAICs.ipynb TODO IK update the WAIC with 1000 reps run
We also tested a uninformative uniform prior with $\mathit{U}(1,6)$, for (i) all $w_{2n+1}$, $w_{2n+1^*}$, $w_{2n^*}$ and (ii) only for $w_{2n+1^*}$, $w_{2n^*}$, using the above informative prior for $w_{2n+1}$. In both cases the algorithm failed to converge.

For the mutation rate, $\mu$, and aneuploidy rate, $\delta$, we used uninformative uniform priors, $\mu \sim \mathit{U}(10^{-9}\ ,\ 10^{-5})$ and $\delta \sim \mathit{U}(10^{-6} , 10^{-2})$. A wider mutation rate prior, $\mu \sim \mathit{U}(10^{-9}\ ,\ 10^{-3})$, produced similar results.

@@

% TODO IK we are jumping between b c s and w2n* w2n+1 w2n+1*. we should report all models and results either with b c s or with w. you can compute b c s from w before computing the HDIs.
This estimated aneuploidy rate agrees with previous estimates. % TODO refs (9.7*10-5 in Zhu 2014, so it's lower). YR: we need more than one ref; also what is the units of Zho 2014?
The mutation rate corresponds to a mutation target size of $10^{4}$, assuming the mutation rate per base pair is roughly $2\cdot10^{-10}$~\citep{Zhu2014} or $3.3\cdot10^{-10}$~\citep{Lynch2008}.

@@

% Fig Curveball 
\begin{figure}[h]
	\centering
	% generated with growth_curves.ipynb and deg = '39' 
	\includegraphics[width=.95\textwidth]{../figures/evo39_fitness_39deg.pdf} 
	% generated with growth_curves_refined.ipynb
	\includegraphics[width=0.95\textwidth]{../figures/refined_vs_evo39_fitness_39deg.pdf}
\caption{
    \textbf{Fitness estimation from growth curves.}
    % TODO IK caption should explain what each panel shows.
    \textbf{(A-D)} Fitness estimation from growth curves of \euwt\ and \anwt\ at \SI{39}{\celsius}: $\hat w_{2n+1}/w_{2n}$=1.024 (95\% CI: 0.959 - 1.115).
    \textbf{(E-H)} Fitness estimation from growth curves of \anwt\ and \eumt\ at \SI{39}{\celsius}: $\hat w_{2n*}/w_{2n+1}$=1.033 (95\% CI: 1.027 - 1.041).\\
    Growth curves previously described in \citet[Figs. 3C, 4A, and S2]{Yona2012}.
	Fitness estimated from growth curves using \texttt{Curveball}, a method for predicting results of competition experiments from growth curve data~\citep[\href{https://curveball.yoavram.com}{curveball.yoavram.com}]{Ram2019}. See \emph{Models and Methods, Prior distributions} for more details.
} 
\label{fig:growth-curves}
\end{figure}
